---
title: "Mateo's YTD Strava Data"
output: html_document
editor_options: 
  chunk_output_type: console
---
Libraries Used 
library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
library(DT)
```
I am an avid user of the Strava application. It's an application on your phone that allows you to track detailed metrics about all you physical activities. Although the application allows you to track many different types of activities, it only provides roll up data for 3 activities, Cycling, Running and Swimming. This bums me out because two activities I capture on a regular basis are Hiking and my Strength Workouts. My bike activities are well represented. So how to I mash it all up so I can see it all? Strava actually allows you to download (once per week) all of your personal data as a CSV. This contains all the data from all of your activities as 78 different variables such as distance, speed, date etc. This is nice but not ideal for any sort of aggregation or analysis. This  script reads the Strava CSV file (as is) and does some basic clean up before creating a data frame. From there you can easily perform basic analysis and plotting. Currently I am creating 3 YTD plots for the activities I most often record and some summaries for total elevation gain, total distance and the total activity time for each. 

```{r}
#read csv file
activities <- read.csv('activities.csv')

#Create a new data frame with clean variable names and dates. Subset data to YTD by activities. After this you can use cleanactivities to do whatever you like with any of the 78 column variables, distance, elevation_gain, date, description etc.
cleanactivities <- activities %>%
  clean_names() %>% 
  #print(activity_type) %>% 
  mutate(activity_date = as.POSIXct(activity_date, format = "%b %d, %Y,%H:%M:%S %p")) %>%
  #For a short time I was logging hikes as runs so I could see the aggregate data in the app. So here I change them to hikes so my totals are accurate.
  mutate(activity_type = str_replace(activity_type, "Run", "Hike")) %>%
  filter(activity_date >= as.Date('2021-01-01')) %>%
  filter(str_detect(activity_type, 'Hike|Ride|Workout'))

#group distances by activity
  dist <- cleanactivities %>% 
  group_by(activity_type) %>% 
  summarise(distance = round(sum(distance) / 1.609)) 
  #print(dist)
#group elevations by activity  
  elev <- cleanactivities %>% 
  group_by(activity_type) %>% 
  summarise(elevation_gain = round(sum(elevation_gain) * 3.281))
  #print(elev)
#group elapsed time by activity
  dur <- cleanactivities %>% 
  group_by(activity_type) %>% 
  summarise(elapsed_time = round(sum(elapsed_time / 60) / 60))
  #print(dur)
```
  
### Strava variable names 

```{r}
names(cleanactivities)
```

### Plot function

```{r}
#Function plots activities, x = summarized data, metric = miles,hours etc., ylabel is metric name.  
  activityPlot <- function(x, metric, ylabel){
      ggplot(x, aes(activity_type, metric, fill = activity_type)) +
      ggtitle(paste( "Total YTD", ylabel, " =", sum(metric) ) ) +
      theme(plot.title = element_text(size = 20, face = "bold")) +
      geom_col() +
      geom_text(aes(label = metric, y = metric - 10), position = position_stack(vjust = 0.5), size = 10) +
      ylab(ylabel) +
      scale_fill_manual(values=c("#CC6666", "#9999CC", "#66CC99"))
  }

```

### Create plots

```{r}
activityPlot(dist, dist$distance, "Distance In Miles")
activityPlot(elev, elev$elevation_gain, "Elevation Gain In Feet")
activityPlot(dur, dur$elapsed_time, "Elapsed Workout Time In Hours")
```


